---
title: "Project 2"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arules)
library(funModeling)
library(RWeka)
```

### 1. Objectives

The data set we are working with predicts if a patient is likely or not to have a chronic kidney disease based on various attributes that are all based on different tests. The objective of the project is to use association rule mining to identify patterns between different attributes of the data set. If the patterns we find are good enough it might be possible for doctors to skip certain tests on patients which could then hopefully save some time in diagnosing different symptoms. 

### 2. Data set description

The data set has 400 different instances that each contain 24 different attributes and one class attribute that tells if the patient has the disease or not. The attributes are:

| Nr. | Name | Type | Description |
| --- | ---- | ---- | ----------- |
| 1. | Age | numerical | Age in years. |
| 2. | Blood Pressure | numerical | bp in mm/Hg. |
| 3. | Specific Gravity | nominal | sg - (1.005,1.010,1.015,1.020,1.025). |
| 4. | Albumin | nominal | al - (0,1,2,3,4,5). |
| 5. | Sugar | nominal | su - (0,1,2,3,4,5). |



6.Red Blood Cells(nominal) ------------ rbc - (normal,abnormal). 
7.Pus Cell (nominal) ------------------ pc - (normal,abnormal).
8.Pus Cell clumps(nominal) ------------ pcc - (present,notpresent). 
9.Bacteria(nominal) ------------------- ba - (present,notpresent). 
10.Blood Glucose Random(numerical)	--- bgr in mgs/dl. 
11.Blood Urea(numerical)	------------- bu in mgs/dl. 
12.Serum Creatinine(numerical)	------- sc in mgs/dl. 
13.Sodium(numerical) ------------------ sod in mEq/L. 
14.Potassium(numerical)	--------------- pot in mEq/L. 
15.Hemoglobin(numerical) -------------- hemo in gms. 
16.Packed Cell Volume(numerical). ----- pcv. 
17.White Blood Cell Count(numerical) -- wc in cells/cumm. 
18.Red Blood Cell Count(numerical)	--- rc in millions/cmm. 
19.Hypertension(nominal)	------------- htn - (yes,no). 
20.Diabetes Mellitus(nominal)	--------- dm - (yes,no). 
21.Coronary Artery Disease(nominal) --- cad - (yes,no). 
22.Appetite(nominal)	----------------- appet - (good,poor). 
23.Pedal Edema(nominal) --------------- pe - (yes,no).	
24.Anemia(nominal) -------------------- ane - (yes,no). 
25.Class (nominal)	------------------- class - (ckd,notckd).

Describe preprocessing (choices in discretization + decisions to ignore attributes)



R kóði - Data read, discretization, remove some attributes?

```{r data}


chronic <- read.arff("chronic_kidney_disease_full.arff")


#range 2-90
chronic$age <- cut(chronic$age, c(2,10,20,30,40,50,60,70,80,90) )
levels(chronic$age) <- c('2-10','10-20','20-30','30-40','40-50',
                         '50-60','60-70','70-80','80-90','90')  

#range 50-180
chronic$bp <- cut(chronic$bp, 8 )
levels(chronic$bp) <- c('50-80','80-90','90-110','110-130','130-150','150-160','160-170','170-180')  
#range 22-490
chronic$bgr <- cut(chronic$bgr, 50 )
#range 1.5-391
chronic$bu <- cut(chronic$bu, 20 )
#range 0.4-76
chronic$sc <- cut(chronic$sc, 20 )
#range 4.5-163
chronic$sod <- cut(chronic$sod, 50 )
#range 2.5-47
chronic$pot <- cut(chronic$pot, 50 )
#range 3.1-17.8
chronic$hemo <- cut(chronic$hemo, 20 )
#range 9-54
chronic$pcv <- cut(chronic$pcv, 20 )
#range 2200-26400 (wbcc í raun er um 10-15)
chronic$wbcc <- cut(chronic$wbcc, 10 )
levels(chronic$wbcc) <- c('2200-5000','5000-7000','7000-8000','8000-9000','9000-10000',
                          '10000-15000','15000-17000','17000-19000','19000-21000','21000+')  
#range 2.1-8
chronic$rbcc <- cut(chronic$rbcc, 20 )



tData <- as (chronic, "transactions")
#size(head(tData))
#LIST(head(tData, 3))

```

### 3. Rule mining process

Using apriori make lists of rules - Most support, Most confidence, Lift .....
Try out different parameters - different length for rules, different thresholds ... more?


```{r seconddata}


#rules <- apriori(tData, parameter=list(support = 0.005, confidence = 0.8, minlen=2), 
                # appearance = list(rhs=c("class=ckd")))  

#inspect(rules)


```

### 4. Results + Recommendations

Resulting rules: Summary (number of rules, general description), and a selection of those
you would show to a client.

Recommendations: What should the client do because of the rules discovered.



