---
title: "Project 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arules)
library(funModeling)
library(RWeka)
```

### 1. Objectives

Potential benifits of finding rules.

### 2. Data set description

What is the data? 
Describe preprocessing (choices in discretization + decisions to ignore attributes)

R kóði - Data read, discretization, remove some attributes?

```{r data}


chronic <- read.arff("chronic_kidney_disease_full.arff")


#bank.full$age <- cut(bank.full$age, 5)
#bank.full$balance <- cut(bank.full$balance, 5)
#bank.full$day <- cut(bank.full$day, 31)
#bank.full$duration <- cut(bank.full$duration, 5)
#bank.full$campaign <- cut(bank.full$campaign, 3)
#bank.full$pdays <- cut(bank.full$pdays, 5)
#bank.full$previous <- cut(bank.full$previous, 3)

tData <- as (chronic, "transactions")
#size(head(tData))
#LIST(head(tData, 3))


```

### 3. Rule mining process

Using apriori make lists of rules - Most support, Most confidence, Lift .....
Try out different parameters - different length for rules, different thresholds ... more?

```{r seconddata}


rules <- apriori(tData, parameter=list(support = 0.005, confidence = 0.8, minlen=2), appearance = list(rhs=c("")))  

inspect(rules)


```

### 4. Results + Recommendations

Resulting rules: Summary (number of rules, general description), and a selection of those
you would show to a client.

Recommendations: What should the client do because of the rules discovered.



